---
title: "bios992_thesis"
author: "Alice Yu"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
library(flextable)
library(gtsummary)
library(haven)
library(rpart)
library(randomForest)
library(ggplot2)
library(GGally)
library(varImp)
library(pROC)
```

## Read in datasets 
Data sets originate from NHANES during the 2017-March 2020 Pre-Pandemic 

```{r}
demo <- read_xpt("~/bios992/P_DEMO.XPT")
dpq <- read_xpt("~/bios992/P_DPQ.XPT")
slq <- read_xpt("~/bios992/P_SLQ.XPT")

#convert to data frame
demo <- demo %>% data.frame()
dpq <- dpq %>% data.frame()
slq <- slq %>% data.frame()
```

## Clean & Recode Data

```{r}
#merge datasets
dpq_slq <- merge(dpq, slq, by="SEQN")
full <- merge(demo, dpq_slq, by="SEQN")
sub_dat <- full %>% select(-SDDSRVYR, -RIDSTATR, -RIDAGEMN, -DMDYRUSZ, -SIALANG, -SIAPROXY, -SIAINTRP, -FIALANG, -FIAPROXY, -FIAINTRP, -MIALANG, -MIAPROXY, -MIAINTRP, -AIALANGA, -RIDEXMON, -RIDEXPRG)

#Recoding those that 'refused' or 'don't know' as missing (NA)
nhanes_codes <- c(7, 77, 777, 77777, 9, 99, 999, 99999)
recode_dat <- sub_dat %>% mutate(DMDBORN4=ifelse(DMDBORN4 %in% nhanes_codes, NA, DMDBORN4),
                                 DMDEDUC2=ifelse(DMDEDUC2 %in% nhanes_codes, NA, DMDEDUC2),
                                 DMDMARTZ=ifelse(DMDMARTZ %in% nhanes_codes, NA, DMDMARTZ),
                                 SLQ300=ifelse(SLQ300 %in% as.character(nhanes_codes), "", SLQ300),
                                 SLQ310=ifelse(SLQ310 %in% as.character(nhanes_codes), "", SLQ310),
                                 SLQ320=ifelse(SLQ320 %in% as.character(nhanes_codes), "", SLQ320),
                                 SLQ330=ifelse(SLQ330 %in% as.character(nhanes_codes), "", SLQ330),
                                 SLQ030=ifelse(SLQ030 %in% nhanes_codes, NA, SLQ030),
                                 SLQ040=ifelse(SLQ040 %in% nhanes_codes, NA, SLQ040),
                                 SLQ050=ifelse(SLQ050 %in% nhanes_codes, NA, SLQ050),
                                 SLQ120=ifelse(SLQ120 %in% nhanes_codes, NA, SLQ120),
                                 DPQ010=ifelse(DPQ010 %in% nhanes_codes, NA, DPQ010),
                                 DPQ020=ifelse(DPQ020 %in% nhanes_codes, NA, DPQ020),
                                 DPQ030=ifelse(DPQ030 %in% nhanes_codes, NA, DPQ030),
                                 DPQ040=ifelse(DPQ040 %in% nhanes_codes, NA, DPQ040),
                                 DPQ050=ifelse(DPQ050 %in% nhanes_codes, NA, DPQ050),
                                 DPQ060=ifelse(DPQ060 %in% nhanes_codes, NA, DPQ060),
                                 DPQ070=ifelse(DPQ070 %in% nhanes_codes, NA, DPQ070),
                                 DPQ080=ifelse(DPQ080 %in% nhanes_codes, NA, DPQ080),
                                 DPQ090=ifelse(DPQ090 %in% nhanes_codes, NA, DPQ090),
                                 DPQ100=ifelse(DPQ100 %in% nhanes_codes, NA, DPQ100))


#changing time to a continuous variable from categorical (convert into hours)
recode_dat <-
  recode_dat %>% separate(
    SLQ300,
    sep = ":",
    into = c("hour", "min"),
    fill = "left",
    convert = T
  ) %>% mutate(min = ifelse(is.na(min), 0, min),
               slq300_hours = (min / 60) + hour) %>% select(-hour,-min)

recode_dat <-
  recode_dat %>% separate(
    SLQ310,
    sep = ":",
    into = c("hour", "min"),
    fill = "left",
    convert = T
  ) %>% mutate(min = ifelse(is.na(min), 0, min),
               slq310_hours = (min / 60) + hour) %>% select(-hour,-min)

recode_dat <-
  recode_dat %>% separate(
    SLQ320,
    sep = ":",
    into = c("hour", "min"),
    fill = "left",
    convert = T
  ) %>% mutate(min = ifelse(is.na(min), 0, min),
               slq320_hours = (min / 60) + hour) %>% select(-hour,-min)

recode_dat <-
  recode_dat %>% separate(
    SLQ330,
    sep = ":",
    into = c("hour", "min"),
    fill = "left",
    convert = T
  ) %>% mutate(min = ifelse(is.na(min), 0, min),
               slq330_hours = (min / 60) + hour) %>% select(-hour,-min)

#adding value labels
recode_dat$RIAGENDR <-
  factor(recode_dat$RIAGENDR,
         levels = c(1, 2),
         labels = c("Male", "Female"))
recode_dat$RIDRETH1 <-
  factor(
    recode_dat$RIDRETH1,
    levels = c(1, 2, 3, 4, 5),
    labels = c(
      "Mexican American",
      "Other Hispanic",
      "Non-Hispanic White",
      "Non-Hispanic Black",
      "Other Race - Including Multi-Racial"
    )
  )
recode_dat$RIDRETH3 <-
  factor(
    recode_dat$RIDRETH3,
    levels = c(1, 2, 3, 4, 6, 7),
    labels = c(
      "Mexican American",
      "Other Hispanic",
      "Non-Hispanic White",
      "Non-Hispanic Black",
      "Non-Hispanic Asian",
      "Other Race - Including Multi-Racial"
    )
  )
recode_dat$DMDBORN4 <-
  factor(
    recode_dat$DMDBORN4,
    levels = c(1, 2),
    labels = c("Born in 50 US states or Washington, DC", "Others")
  )
recode_dat$DMDEDUC2 <-
  factor(
    recode_dat$DMDEDUC2,
    levels = c(1, 2, 3, 4, 5),
    labels = c(
      "Less than 9th grade",
      "9-11th grade (Includes 12th grade, no diploma)",
      "High school graduate/GED or equivalent",
      "Some college or AA degree",
      "College graduate or above"
    )
  )
recode_dat$DMDMARTZ <-
  factor(
    recode_dat$DMDMARTZ,
    levels = c(1, 2, 3),
    labels = c(
      "Married/Living with partner",
      "Widowed/Divorced/Separated",
      "Never married"
    )
  )
recode_dat$SLQ030 <-
  factor(
    recode_dat$SLQ030,
    levels = c(0, 1, 2, 3),
    labels = c(
      "Never",
      "Rarely: 1-2 Nights a week",
      "Occasionally: 3-4 nights a week",
      "Frequently: 5 or more nights a week"
    )
  )
recode_dat$SLQ040 <-
  factor(
    recode_dat$SLQ040,
    levels = c(0, 1, 2, 3),
    labels = c(
      "Never",
      "Rarely: 1-2 Nights a week",
      "Occasionally: 3-4 nights a week",
      "Frequently: 5 or more nights a week"
    )
  )
recode_dat$SLQ050 <-
  factor(recode_dat$SLQ050,
         levels = c(1, 2),
         labels = c("Yes", "No"))
recode_dat$SLQ120 <-
  factor(
    recode_dat$SLQ120,
    levels = c(0, 1, 2, 3, 4),
    labels = c(
      "Never",
      "Rarely: 1 time a month",
      "Sometimes: 2-4 times a month",
      "Often: 5-15 times a month",
      "Almost always: 16-30 times a month"
    )
  )
recode_dat$DPQ010 <-
  factor(
    recode_dat$DPQ010,
    levels = c(0, 1, 2, 3),
    labels = c(
      "Not at all",
      "Several days",
      "More than half the days",
      "Nearly every day"
    )
  )
recode_dat$DPQ020 <-
  factor(
    recode_dat$DPQ020,
    levels = c(0, 1, 2, 3),
    labels = c(
      "Not at all",
      "Several days",
      "More than half the days",
      "Nearly every day"
    )
  )
recode_dat$DPQ030 <-
  factor(
    recode_dat$DPQ030,
    levels = c(0, 1, 2, 3),
    labels = c(
      "Not at all",
      "Several days",
      "More than half the days",
      "Nearly every day"
    )
  )
recode_dat$DPQ040 <-
  factor(
    recode_dat$DPQ040,
    levels = c(0, 1, 2, 3),
    labels = c(
      "Not at all",
      "Several days",
      "More than half the days",
      "Nearly every day"
    )
  )
recode_dat$DPQ050 <-
  factor(
    recode_dat$DPQ050,
    levels = c(0, 1, 2, 3),
    labels = c(
      "Not at all",
      "Several days",
      "More than half the days",
      "Nearly every day"
    )
  )
recode_dat$DPQ060 <-
  factor(
    recode_dat$DPQ060,
    levels = c(0, 1, 2, 3),
    labels = c(
      "Not at all",
      "Several days",
      "More than half the days",
      "Nearly every day"
    )
  )
recode_dat$DPQ070 <-
  factor(
    recode_dat$DPQ070,
    levels = c(0, 1, 2, 3),
    labels = c(
      "Not at all",
      "Several days",
      "More than half the days",
      "Nearly every day"
    )
  )
recode_dat$DPQ080 <-
  factor(
    recode_dat$DPQ080,
    levels = c(0, 1, 2, 3),
    labels = c(
      "Not at all",
      "Several days",
      "More than half the days",
      "Nearly every day"
    )
  )
recode_dat$DPQ090 <-
  factor(
    recode_dat$DPQ090,
    levels = c(0, 1, 2, 3),
    labels = c(
      "Not at all",
      "Several days",
      "More than half the days",
      "Nearly every day"
    )
  )
recode_dat$DPQ100 <-
  factor(
    recode_dat$DPQ100,
    levels = c(0, 1, 2, 3),
    labels = c(
      "Not at all difficult",
      "Somewhat difficult",
      "Very difficult",
      "Extremely difficult"
    )
  )


#changing categorical variables that use dummy codes to be type character from numeric
recode_dat <- recode_dat %>% mutate(DMDBORN4=as.character(DMDBORN4),
                                    RIAGENDR=as.character(RIAGENDR),
                                    RIDRETH1=as.character(RIDRETH1),
                                    RIDRETH3=as.character(RIDRETH3),
                                    DMDEDUC2=as.character(DMDEDUC2),
                                    DMDMARTZ=as.character(DMDMARTZ),
                                    SLQ030=as.character(SLQ030),
                                    SLQ040=as.character(SLQ040),
                                    SLQ050=as.character(SLQ050),
                                    SLQ120=as.character(SLQ120),
                                    DPQ010=as.character(DPQ010),
                                    DPQ020=as.character(DPQ020),
                                    DPQ030=as.character(DPQ030),
                                    DPQ040=as.character(DPQ040),
                                    DPQ050=as.character(DPQ050),
                                    DPQ060=as.character(DPQ060),
                                    DPQ070=as.character(DPQ070),
                                    DPQ080=as.character(DPQ080),
                                    DPQ090=as.character(DPQ090),
                                    DPQ100=as.character(DPQ100)
                                    )


```

## Dealing with Missing data
Imputing missing categorical data with mode and continuous data with mean 

```{r}
#comparing number of observations of data set with missing data vs dropping observations with >= 1 missing variable
no_na <- recode_dat %>% drop_na()

#identifying and evaluating # of missing variables and values 
missing <- colSums(is.na(recode_dat))
missing_frame <- data.frame(t(missing))
missing_vars <- missing_frame[,missing_frame>0]

#function to find the mode for categorical variables 
getMode <- function(data) {
  data=data[nchar(as.character(data))>0]
  unique_v <- unique(data)
  unique_v[which.max(tabulate(match(data, unique_v)))]
  
}

#imputing missing values
imputed_dat <- recode_dat #replicating dataset

for(cols in colnames(imputed_dat)) {
  if (cols %in% names(imputed_dat[, sapply(imputed_dat, is.numeric)])) {
    imputed_dat <-
      imputed_dat %>% mutate(!!cols := replace(
        !!rlang::sym(cols),
        is.na(!!rlang::sym(cols)),
        mean(!!rlang::sym(cols), na.rm = TRUE)
      ))
  } else {
    imputed_dat <-
      imputed_dat %>% mutate(!!cols := replace(
        !!rlang::sym(cols),
        !!rlang::sym(cols) == "" |
          is.na(!!rlang::sym(cols)),
        getMode(!!rlang::sym(cols))
      ))
  }
}

#check if there still any missing values
missing <- colSums(is.na(imputed_dat))
```

## Data Exploration
Looking at the demographic and characteristics of participants 

```{r}
#create summary statistics table 
recode_dat2 <- recode_dat %>% drop_na()
recode_dat2$type <- "missing"
imputed_dat$type <- "non-missing"

merged_dat <- rbind(recode_dat2,imputed_dat)


merged_dat %>%
  select(-SEQN,-WTINTPRP,-WTMECPRP,-SDMVPSU,-SDMVSTRA) %>%
  tbl_summary(
    by = type,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = all_continuous() ~ 2,
    missing_text = "(Missing)",
    label = c(
      RIAGENDR ~ "Sex",
      RIDAGEYR ~ "Age (years)",
      RIDRETH1 ~ "Race/Hispanic Origin",
      RIDRETH3 ~ "Race/Hispanic origin with NH Asian",
      DMDBORN4 ~ "Country of birth",
      DMDEDUC2 ~ "Education level (Adults 20+)",
      DMDMARTZ ~ "Marital status",
      INDFMPIR ~ "Ratio of family income to poverty",
      SLD012 ~ "Sleep hours (weekdays/workdays)",
      SLD013 ~ "Sleep hours (weekends)",
      SLQ030 ~ "Snore Amount",
      SLQ040 ~ "Frequency of snorting or stop breathing",
      SLQ050 ~ "Told doctor trouble sleeping",
      SLQ120 ~ "Frequency of feeling overly sleepy during day",
      DPQ010 ~ "Have little interest in doing things",
      DPQ020 ~ "Feeling down, depressed or hopeless",
      DPQ030 ~ "Trouble sleeping or sleeping too much",
      DPQ040 ~ "Feeling tired or having little energy",
      DPQ050 ~ "Poor appetite or overeating",
      DPQ060 ~ "Feeling bad about yourself",
      DPQ070 ~ "Trouble concentrating on things",
      DPQ080 ~ "Moving or speaking slowly or too fast",
      DPQ090 ~ "Thoughts you would be better off dead",
      DPQ100 ~ "Difficulty these problems have caused",
      slq300_hours ~ "Sleep time on weekdays/workdays",
      slq310_hours ~ "Wake time on weekdays/workdays",
      slq320_hours ~ "Sleep time on weekends",
      slq330_hours ~ "Wake time on weekends"
    )
  ) %>%
  modify_header(label = "****") %>%
  bold_labels() %>%
  as_flex_table() %>%
  bold(part = "header") %>%
  font(fontname = 'Computer Modern')


```

```{r}
#classifying depression using PHQ-9 depression screening instrument

#create copy of dataset and change depression questions back to continuous so can calculate depression score
to_numeric <- function(var) {
  if(var=="Not at all") {
    var <- 0
  } else if (var=="Several days") {
    var <- 1
  } else if (var=="More than half the days") {
    var <- 2
  } else {
    var <- 3
  }
 
}

analysis_dat <- imputed_dat

#converting depression questions back to numeric
for(i in 1:nrow(analysis_dat)) {
  analysis_dat$DPQ010_num[i] <- to_numeric(analysis_dat$DPQ010[i])
}

for(i in 1:nrow(analysis_dat)) {
  analysis_dat$DPQ020_num[i] <- to_numeric(analysis_dat$DPQ020[i])
}

for(i in 1:nrow(analysis_dat)) {
  analysis_dat$DPQ030_num[i] <- to_numeric(analysis_dat$DPQ030[i])
}

for(i in 1:nrow(analysis_dat)) {
  analysis_dat$DPQ040_num[i] <- to_numeric(analysis_dat$DPQ040[i])
}

for(i in 1:nrow(analysis_dat)) {
  analysis_dat$DPQ050_num[i] <- to_numeric(analysis_dat$DPQ050[i])
}

for(i in 1:nrow(analysis_dat)) {
  analysis_dat$DPQ060_num[i] <- to_numeric(analysis_dat$DPQ060[i])
}

for(i in 1:nrow(analysis_dat)) {
  analysis_dat$DPQ070_num[i] <- to_numeric(analysis_dat$DPQ070[i])
}

for(i in 1:nrow(analysis_dat)) {
  analysis_dat$DPQ080_num[i] <- to_numeric(analysis_dat$DPQ080[i])
}

for(i in 1:nrow(analysis_dat)) {
  analysis_dat$DPQ090_num[i] <- to_numeric(analysis_dat$DPQ090[i])
}


#calculating depression score
analysis_dat <- analysis_dat %>% select(-type)
analysis_dat <- analysis_dat %>% mutate(depression_score=rowSums(.[34:42]))

#categorizing depression, based on PHQ-9 depression instrument
for(i in 1:nrow(analysis_dat)) {
  if(analysis_dat$depression_score[i] >= 0 & analysis_dat$depression_score[i] <= 4) {
    analysis_dat$depression_status[i] <- 1 #minimal
  } else if(analysis_dat$depression_score[i] >= 5 & analysis_dat$depression_score[i]<=9) {
    analysis_dat$depression_status[i] <- 2 #mild
  } else if(analysis_dat$depression_score[i]>=10 & analysis_dat$depression_score[i]<=14) {
    analysis_dat$depression_status[i] <- 3 #moderate
  } else if(analysis_dat$depression_score[i]>=15 & analysis_dat$depression_score[i]<=19) {
    analysis_dat$depression_status[i] <- 4 #moderately severe
  } else {
    analysis_dat$depression_status[i] <- 5 #severe
  }
}


#group by sleep disorder vs none  
#Classification of sleep disorder:
#https://www-sciencedirect-com.libproxy.lib.unc.edu/science/article/pii/S016503272101209X

#SLD012 - sleep duration during weekdays  
  #insufficient (sleepHr_status = 1) <6
  #normal (sleepHr_status = 2) 6-8
  #excessive (sleepHr_status = 3) >8
#SLQ050 - Told doctor trouble sleeping
#sleep debt: difference b/w weekdays and weekends (>=2 is bad)

#categorizing sleep duration
analysis_dat$sleepHr_status <-
  ifelse(
    analysis_dat$SLD012 < 6,
    1,
    ifelse(analysis_dat$SLD012 <= 6 & analysis_dat$SLD012 >= 8, 2, 3)
  ) 

#converting back to numeric for variable SLQ050
for(i in 1:nrow(analysis_dat)) {
  if(analysis_dat$SLQ050[i]=="Yes") {
    analysis_dat$SLQ050_num[i] <- 1
  } else {
    analysis_dat$SLQ050_num[i] <- 2
  }
}

#creating variable sleep_debt
analysis_dat <- analysis_dat %>% mutate(sleep_debt=round((SLD013-SLD012),3))

#categorizing sleep disorder, 0=no, 1=yes
for(i in 1:nrow(analysis_dat)) {
  if((analysis_dat$sleepHr_status[i] %in% c(1,3) & analysis_dat$SLQ050_num[i]==1) | analysis_dat$sleep_debt[i]>=2) {
    analysis_dat$sleep_disorder[i] <- 1
  } else {
    analysis_dat$sleep_disorder[i] <- 0
  }
}

#subsetting to only those suffering with sleep disorder
sleepDisorder_dat <- analysis_dat[analysis_dat$sleep_disorder==1,]

#creating dependent variable into binary 
#0=no depression, 1=depression
sleepDisorder_dat$outcome <- ifelse(sleepDisorder_dat$depression_score<=9,0,1) 

```

```{r}
#descriptive statistics of sleep disorder data set
sleepDisorder_dat %>%
  select(
    -SEQN,
    -WTINTPRP,
    -WTMECPRP,
    -SDMVPSU,
    -SDMVSTRA,
    -DPQ010_num,
    -DPQ020_num,
    -DPQ030_num,
    -DPQ040_num,
    -DPQ050_num,
    -DPQ060_num,
    -DPQ070_num,
    -DPQ080_num,
    -DPQ090_num,
    -sleepHr_status,
    -SLQ050_num,
    -sleep_disorder,
    -depression_status
  ) %>%
  tbl_summary(
    by = outcome,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = all_continuous() ~ 2,
    missing_text = "(Missing)",
    label = c(
      RIAGENDR ~ "Gender",
      RIDAGEYR ~ "Age (years)",
      RIDRETH1 ~ "Race/Hispanic Origin",
      RIDRETH3 ~ "Race/Hispanic origin with NH Asian",
      DMDBORN4 ~ "Country of birth",
      DMDEDUC2 ~ "Education level (Adults 20+)",
      DMDMARTZ ~ "Marital status",
      INDFMPIR ~ "Ratio of family income to poverty",
      SLD012 ~ "Sleep hours (weekdays/workdays)",
      SLD013 ~ "Sleep hours (weekends)",
      SLQ030 ~ "Snore Amount",
      SLQ040 ~ "Frequency of snorting or stop breathing",
      SLQ050 ~ "Told doctor trouble sleeping",
      SLQ120 ~ "Frequency of feeling overly sleepy during day",
      DPQ010 ~ "Have little interest in doing things",
      DPQ020 ~ "Feeling down, depressed or hopeless",
      DPQ030 ~ "Trouble sleeping or sleeping too much",
      DPQ040 ~ "Feeling tired or having little energy",
      DPQ050 ~ "Poor appetite or overeating",
      DPQ060 ~ "Feeling bad about yourself",
      DPQ070 ~ "Trouble concentrating on things",
      DPQ080 ~ "Moving or speaking slowly or too fast",
      DPQ090 ~ "Thoughts you would be better off dead",
      DPQ100 ~ "Difficulty these problems have caused",
      slq300_hours ~ "Sleep time on weekdays/workdays",
      slq310_hours ~ "Wake time on weekdays/workdays",
      slq320_hours ~ "Sleep time on weekends",
      slq330_hours ~ "Wake time on weekends",
      depression_score ~ "Depression Score",
      sleep_debt ~ "Sleep Debt (sleep time on weekends - weekdays)"
    )
  ) %>%
  modify_header(label = "****",
                stat_1 = "**No Depression**, N={n}",
                stat_2 = "**Depression**,N={n}") %>%
  bold_labels() %>%
  as_flex_table() %>%
  bold(part = "header") %>%
  font(fontname = 'Computer Modern')


```

```{r, fig.width=12, fig.height=8}
#look at correlations, relationships b/w sleep disorder and depression variables 
ggpairs(
  sleepDisorder_dat,
  columns = c(
    'SLD012',
    'SLD013',
    'slq300_hours',
    'slq310_hours',
    'slq320_hours',
    'slq330_hours',
    'depression_score',
    'sleep_debt'
  ),
  ggplot2::aes(color = as.character(outcome))
) 

```

## Using Machine Learning for Predictions

```{r}
#subsetting dataset to only keep variables of interest 
sleep_dat <-
  sleepDisorder_dat %>% select(
    RIAGENDR,
    RIDAGEYR,
    RIDRETH3,
    DMDEDUC2,
    DMDMARTZ,
    WTMECPRP,
    INDFMPIR,
    SLD012,
    SLD013,
    SLQ030,
    SLQ040,
    SLQ050,
    SLQ120,
    slq300_hours,
    slq310_hours,
    slq320_hours,
    slq300_hours,
    slq330_hours,
    sleep_debt,
    outcome
  )

#check balanced or imbalanced
no_depression_num <- sum(sleep_dat$outcome == 0)
depression_num <- sum(sleep_dat$outcome == 1)

no_depression_num/depression_num

```

### CART without Pruning 
Not using pruning due to imbalanced dataset
```{r}
#for reproducibility purposes
set.seed(20)

#create 5-fold Cross validation (CV)
tt_indicies <- createFolds(y=sleep_dat$outcome,k=5)

perF_conMatrix <- list()
perF_roc_obs <- list()

for(i in 1:length(tt_indicies)) {
  #creating train and test sets
  sleep_train <- sleep_dat[-tt_indicies[[i]],]
  sleep_test <- sleep_dat[tt_indicies[[i]],]
  
  #fit tree
  #WTMECPRP use this variable for sampling weights b/c sleep disorder and depression are from examination   portion (MEC)
  cart <-
    rpart(
      outcome ~ RIAGENDR + RIDAGEYR + RIDRETH3 + DMDEDUC2 + DMDMARTZ + INDFMPIR +
        SLD012 + SLD013 + SLQ030 + SLQ040 + SLQ050 + SLQ120 + slq300_hours + slq310_hours +
        slq320_hours + slq330_hours + sleep_debt,
      sleep_train,
      weight = WTMECPRP
    )
  
  #predict on test set 
  sleep_test$cart_predict <- predict(cart,newdata=sleep_test)
  
  #add in test set predictions as binary factor
  sleep_test <-
    sleep_test %>% mutate(predict_outcome = factor(ifelse(cart_predict < 0.5, "No", "Yes")))
  
  #Creating confusion matrix
  sleep_test$depression <- ifelse(sleep_test$outcome == 0, "No","Yes")
  depression_outcome <- as.factor(sleep_test$depression)
  conMatrix <- confusionMatrix(data=sleep_test$predict_outcome,reference=depression_outcome,positive="Yes")
  
  perF_conMatrix[[i]] <- conMatrix 
  
  #add ROC
  perF_roc_obs[[i]] <- roc(response=sleep_test$outcome,predictor=sleep_test$cart_predict)
  
}

#overall ROC and AUC
responses <- list()
predictors <- list()

for (i in 1:length(perF_roc_obs)) {
  responses[[i]] <- perF_roc_obs[[i]]$response
  predictors[[i]] <- perF_roc_obs[[i]]$predictor
}

response_list <- do.call("c", responses)
predictors_list <- do.call("c", predictors)

roc_dat <-
  data.frame(response = response_list, predictor = predictors_list)


roc <- roc(roc_dat,
           response,
           predictor,
           plot = TRUE,
           print.auc = TRUE,
           main="ROC curve with overall AUC")

#average recall and precision
precision_sum <- 0 
recall_sum <- 0

for(i in 1:length(perF_conMatrix)) {
  precision_sum <- precision_sum + perF_conMatrix[[i]]$byClass[5]
  recall_sum <- recall_sum + perF_conMatrix[[i]]$byClass[6]
}

cart_precision_mean <- precision_sum/length(perF_conMatrix)
cart_recall_mean <- recall_sum/length(perF_conMatrix)

```

### Random Forest and Feature Importance
Use error rate to select tuning parameter instead of MSE because of categorical outcome
```{r}
set.seed(20)

imp_score <- list()
perF_conMatrix <- list()
perF_roc_obs <- list()

for(i in 1:length(tt_indicies)) {
  sleep_train <- sleep_dat[-tt_indicies[[i]],]
  sleep_test <- sleep_dat[tt_indicies[[i]],]
  
  #outcome to factor
  sleep_train$outcome <- as.factor(sleep_train$outcome)
  
  #fit random forest
  reg_rf <-
    randomForest(
      outcome ~ RIAGENDR + RIDAGEYR + RIDRETH3 + DMDEDUC2 + DMDMARTZ + INDFMPIR +
        SLD012 + SLD013 + SLQ030 + SLQ040 + SLQ050 + SLQ120 + slq300_hours + slq310_hours +
        slq320_hours + slq330_hours + sleep_debt,
      sleep_train,
      importance = TRUE,
      weights = WTMECPRP
    )
  
  #try different parameter values for trees and number of features (m)
  rf_err <- list() #random forest error rate
  p <- ncol(sleep_train)-1 #number of predictors
  trees <- c(50, 250, 500)
  m <- c(p/2,sqrt(p))
  count <- 1

  for(j in 1:length(trees)) {
    for(k in 1:length(m)) {
      rf_pred_tune <-
        randomForest(
          outcome ~ RIAGENDR + RIDAGEYR + RIDRETH3 + DMDEDUC2 + DMDMARTZ + INDFMPIR +
            SLD012 + SLD013 + SLQ030 + SLQ040 + SLQ050 + SLQ120 + slq300_hours + slq310_hours +
            slq320_hours + slq330_hours + sleep_debt,
          sleep_train,
          ntree = trees[j],
          mtry = m[k],
          weights = WTMECPRP
        )
      rf_err[[count]] <-
        data.frame("ntree" = trees[j],
                   "pred_no" = m[k],
                   "oob_error" = rf_pred_tune$err.rate[trees[j],1])

      count <- count + 1
    }
  }
  
  rf_err_df <- do.call("rbind", rf_err)
  
  #refit on training using best no. of predictors at split
  best_error <- which(rf_err_df$oob_error == min(rf_err_df$oob_error))
  
  #if multiple best errors, i.e. same oob error
  best_error_num <- ifelse(length(best_error)>1, best_error[1],best_error)
  
  print(i)
  reg_rf <- randomForest(
    outcome ~ RIAGENDR + RIDAGEYR + RIDRETH3 + DMDEDUC2 + DMDMARTZ + INDFMPIR +
      SLD012 + SLD013 + SLQ030 + SLQ040 + SLQ050 + SLQ120 + slq300_hours + slq310_hours +
      slq320_hours + slq330_hours + sleep_debt,
    sleep_train,
    ntree = rf_err_df$ntree[best_error_num], 
    mtry = rf_err_df$pred_no[best_error_num],
    importance = TRUE,
    weights = WTMECPRP
  )
  
  #predict on test set
  sleep_test$pred_rf <- predict(reg_rf,newdata=sleep_test)
  
  #create confusion matrix
  perF_conMatrix[[i]] <- confusionMatrix(as.factor(sleep_test$pred_rf), as.factor(sleep_test$outcome)) 
  
  #add ROC
  perF_roc_obs[[i]] <- roc(response=as.numeric(sleep_test$outcome),predictor=as.numeric(sleep_test$pred_rf))
  
  
  #saving feature importance from RF model
  imp_score[[i]] <- caret::varImp(reg_rf)
  imp_score[[i]] <- imp_score[[i]] %>% rownames_to_column("var")
  imp_score[[i]]$var <- imp_score[[i]]$var %>% as.factor()
  
}

#overall ROC and AUC
responses <- list()
predictors <- list()

for (i in 1:length(perF_roc_obs)) {
  responses[[i]] <- perF_roc_obs[[i]]$response
  predictors[[i]] <- perF_roc_obs[[i]]$predictor
}

response_list <- do.call("c", responses)
predictors_list <- do.call("c", predictors)

roc_dat <-
  data.frame(response = response_list, predictor = predictors_list)


roc <- roc(roc_dat,
           response,
           predictor,
           plot = TRUE,
           print.auc = TRUE,
           main="ROC curve with overall AUC") 

#average recall and precision
precision_sum <- 0 
recall_sum <- 0

for(i in 1:length(perF_conMatrix)) {
  precision_sum <- precision_sum + perF_conMatrix[[i]]$byClass[5]
  recall_sum <- recall_sum + perF_conMatrix[[i]]$byClass[6]
}

rf_precision_mean <- precision_sum/length(perF_conMatrix)
rf_recall_mean <- recall_sum/length(perF_conMatrix)


#table of feature importance by fold and average
imp_score_sub <- list()

for(i in 1:length(imp_score)) {
  imp_score_sub[[i]] <- subset(imp_score[[i]],select=-c(3))
  colnames(imp_score_sub[[i]]) <- c("var",paste0("Fold",i))
}

imp_feat <- Reduce(function(...) merge(...,all=TRUE,by="var"),imp_score_sub)
imp_feat$Average <- rowMeans(imp_feat[2:6])
imp_feat <- imp_feat[,c(1,7,2,3,4,5,6)]

imp_feat %>% flextable() %>% font(fontname="Computer Modern")

#barplot for comparing variables importance 
imp_bar <-
  ggplot(data = imp_feat) + geom_bar(
    stat = "identity",
    mapping = aes(
      x = Average,
      y = reorder(var, Average),
      fill = var
    ),
    show.legend = FALSE,
    width = 1
  ) + labs(x = "Average Variable Importance", y = "Predictor") + theme(panel.background = element_blank())

imp_bar

```

### Comparing model results
Did not use accuracy due to imbalanced data.
```{r}
avg <-
  data.frame(
    c("Random Forest", "CART"),
    c(rf_precision_mean, cart_precision_mean),
    c(rf_recall_mean, cart_recall_mean)
  )
colnames(avg) <- c("var","Average Precision","Average Recall")

avg %>% flextable() %>% font(fontname = "Computer Modern") %>% set_header_labels(
  var = "",
  `Average Precision` = "Precision",
  `Average Recall` = "Recall"
)

```
